Theories of deliberated choice

### 2023-06-18 simple.tex
Rm note about reverse impl
Notation f

We now say that $C$ is falsifiable iff $\forall {\allleadstog} \in P^{I × \allhist}: ¬C({\allleadstog}) \implies \exists O \subseteq {\allleadstog} \suchthat O \implies ¬C$, meaning that
$\forall {\allleadstog} \notin C: \exists O \subseteq {\allleadstog} \suchthat S_O \cap C = \emptyset$.
\begin{example}
	$(\ar) \ileadstog \phi \lor  (\ar, \ar) \ileadstog \phi \lor …$ is falsifiable.
	Consider any ${\allleadstog} \notin C$, thus satisfying $\exists k \suchthat (\ar)^k \nileadstog \phi$. Consider $O = \set{(i, (\ar)^k, {\allleadstog}(i, (\ar)^k) ≠ \phi)}$, thus with $O \subset {\allleadstog}$. We have $S_O \cap C = \emptyset$: consider any $O \subset f$, it is not in $C$.
\end{example}
\begin{example}
	\label{ex:unfals}
	$(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi$ is not falsifiable.
	Consider any ${\allleadstog} \notin C$, thus satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
	Consider any $O \subseteq {\allleadstog}$ and let us prove that $S_O \cap C ≠ \emptyset$.
	By definition of a general observation, $O$ may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $O$, pick any $f$ such that $O \subset {f} \land t \in {f}$. We have ${f} \in S_O \cap C ≠ \emptyset$.
\end{example}
\begin{example}
	$\hist \ileadstog[j] \phi \land [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable.
	Consider any $\allleadstog$ satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$, thus with ${\allleadstog} \notin C$.
	\Cref{ex:unfals} concludes.
\end{example}

Given an exposure protocol $\allleadsto$, deliberated choices are represented by an element $\pi({\allleadsto}) \subseteq I × P$ which is right unique (meaning that $\forall i \in I: \card{\pi({\allleadsto})(i)} ≤ 1$).
Equivalently, let $(P \cup \set{\emptyset})^I$ denote the right unique relations over $I × P$; and then the deliberated choices are an element $\pi({\allleadsto}) \in (P \cup \set{\emptyset})^I$.
We say that $C$ is informative iff $\exists {\allleadstog} \in (P \cup \set{\emptyset})^I \suchthat C \implies ¬{\allleadstog}$, meaning that $\pi(C) = \set{\pi(\allleadstog) \suchthat {\allleadstog} \in C} ≠ (P \cup \set{\emptyset})^I$.

Given an exposure protocol $\allleadsto$, we say that a claim $C$ holds iff $C \ni {\allleadsto}$.
$C$ is $\allleadsto$-suitable iff it is informative and falsifiable and ${\allleadsto} \in C$.

The following theorem verifies that when $C$ holds, it will not be falsified ex post (though it may be (a priori) falsifiable).
\begin{theorem}
	\label{th:ex-post}
	$\forall C \ni {\allleadsto}: \nexists O \subseteq {\allleadsto} \suchthat S_O \cap C = \emptyset$.
\end{theorem}
\begin{proof}
	We have, $\forall C \suchthat C \ni {\allleadsto}, \forall O \subseteq {\allleadsto}$: $O \subseteq {\allleadsto} \in C$, equivalently, ${\allleadsto} \in S_O \cap C$, thus $S_O \cap C ≠ \emptyset$.
\end{proof}
Therefore, falsifiability guarantees that a claim holds if, and only if, it is not a posteriori falsifiable.
\begin{theorem}
	\label{th:holdsiff}
	For all falsifiable $C$: ${\allleadsto} \in C$ iff $\nexists O \subseteq {\allleadsto} \suchthat S_O \cap C = \emptyset$.
\end{theorem}
\begin{proof}
	Consider any falsifiable $C$.
	By definition of falsifiability, $\forall {\allleadstog}: \nexists O \subseteq {\allleadstog} \suchthat S_O \cap C = \emptyset \implies {\allleadstog} \in C$.
	In particular,
	$\nexists O \subseteq {\allleadsto} \suchthat S_O \cap C = \emptyset \implies {\allleadsto} \in C$.
	The reverse implication is proven by \cref{th:ex-post}.
\end{proof}
Note that the reverse implication of $\cref{th:holdsiff} does not hold: a claim may hold and be not a posteriori falsifiable but be not a priori falsifiable.

\begin{example}[To be corrected]
Consider any $\allleadsto$ such that $\exists \ar, \phi \suchthat \forall i: \ar \ileadstost \phi$.
Then some suitable $C$ exists, for example, $\forall i: \ar \ileadstost \phi$.
Note that it is falsifiable but not verifiable (that is, no $O$ implies $C$).
\end{example}

>>>>

We now say that $C$ is \emph{falsifiable} iff $\forall {\allleadstog} \in P^{I × \allhist}: ¬C({\allleadstog}) \implies \exists O \subseteq {\allleadstog} \suchthat O \implies ¬C$, meaning that
$\forall {\allleadstog} \notin C: \exists O \subseteq {\allleadstog} \suchthat S_O \cap C = \emptyset$.
\begin{example}
	$(\ar) \ileadstog \phi \lor  (\ar, \ar) \ileadstog \phi \lor …$ is falsifiable.
	Consider any ${\allleadstog} \notin C$, thus satisfying $\exists k \suchthat (\ar)^k \nileadstog \phi$. Consider $O = \set{(i, (\ar)^k, {\allleadstog}(i, (\ar)^k) ≠ \phi)}$, thus with $O \subset {\allleadstog}$. We have $S_O \cap C = \emptyset$: consider any $O \subset f$, it is not in $C$.
\end{example}
\begin{example}
	\label{ex:unfals}
	$(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi$ is not falsifiable.
	Consider any ${\allleadstog} \notin C$, thus satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
	Consider any $O \subseteq {\allleadstog}$ and let us prove that $S_O \cap C ≠ \emptyset$.
	By definition of a general observation, $O$ may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $O$, pick any ${\allleadstog'}$ such that $O \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have ${\allleadstog'} \in S_O \cap C ≠ \emptyset$.
\end{example}
\begin{example}
	$\hist \ileadstog[j] \phi \land [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable.
	Consider any $\allleadstog$ satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$, thus with ${\allleadstog} \notin C$.
	\Cref{ex:unfals} concludes.
\end{example}
\begin{example}
	Consider any $\ar$ and $\phi$. The claim $\forall i: \ar \ileadstostg \phi$ is falsifiable.
	Given any ${\allleadstog} \notin C$, consider any $i \suchthat \ar \nileadstostg \phi$ and any $\hist \suchthat \hist \nileadstog \phi \land \ar \in \histend$, and define $O = \set{(i, \hist, {\allleadstog}(i, \hist) ≠ \phi)}$.
\end{example}

Given an exposure protocol $\allleadsto$, we say that a claim $C$ \emph{holds} iff $C \ni {\allleadsto}$.

The following theorem verifies that when $C$ holds, it will not be falsified ex post (though it may be (a priori) falsifiable).
\begin{theorem}
	\label{th:ex-post}
	$\forall C \ni {\allleadsto}: \nexists O \subseteq {\allleadsto} \suchthat S_O \cap C = \emptyset$.
\end{theorem}
\begin{proof}
	We have, $\forall C \suchthat C \ni {\allleadsto}, \forall O \subseteq {\allleadsto}$: $O \subseteq {\allleadsto} \in C$, equivalently, ${\allleadsto} \in S_O \cap C$, thus $S_O \cap C ≠ \emptyset$.
\end{proof}
Therefore, falsifiability guarantees that a claim holds if, and only if, it is not a posteriori falsifiable. (TODO syntactically fals VS empirically fals.)
\begin{theorem}
	\label{th:holdsiff}
	For all falsifiable $C$: ${\allleadsto} \in C$ iff $\nexists O \subseteq {\allleadsto} \suchthat S_O \cap C = \emptyset$.
\end{theorem}
\begin{proof}
	Consider any falsifiable $C$.
	By definition of falsifiability, $\forall {\allleadstog}: \nexists O \subseteq {\allleadstog} \suchthat S_O \cap C = \emptyset \implies {\allleadstog} \in C$.
	In particular,
	$\nexists O \subseteq {\allleadsto} \suchthat S_O \cap C = \emptyset \implies {\allleadsto} \in C$.
	The reverse implication is proven by \cref{th:ex-post}.
\end{proof}

Given an exposure protocol $\allleadsto$, deliberated choices are represented by an element $\pi({\allleadsto}) \subseteq I × P$ which is right unique (meaning that $\forall i \in I: \card{\pi({\allleadsto})(i)} ≤ 1$).
Equivalently, let $(P \cup \set{\emptyset})^I$ denote the right unique relations over $I × P$; and then the deliberated choices are an element $\pi({\allleadsto}) \in (P \cup \set{\emptyset})^I$.
We say that $C$ is \emph{informative} iff $\exists {\allleadstog} \in (P \cup \set{\emptyset})^I \suchthat C \implies ¬{\allleadstog}$, meaning that $\pi(C) = \set{\pi(\allleadstog) \suchthat {\allleadstog} \in C} ≠ (P \cup \set{\emptyset})^I$.

$C$ is \emph{suitable} iff it holds and is informative and falsifiable.
\begin{example}[To be corrected]
	Assume that $\exists \ar, \phi \suchthat \forall i: \ar \ileadstost \phi$.
	Then $\forall i: \ar \ileadstostg \phi$ is suitable.

	Note that it is falsifiable but not verifiable (that is, no $O$ implies $C$).
\end{example}
